Pattern Avoidance

### 2014-11-07 project2/paper/paper.tex
more attributions
claimed stuff and edited Section 4 some

\emph{Yajit wrote this section. }

@@

\subsection{Flipping}
In this section we introduce the concept of flipping and use generating trees to prove the `Flipping Lemma.'
\begin{definition}
We define the \emph{flip} of a sequence $a$ as the sequence $b$ with the same elements as $a$, but with the largest element swapped with the smallest element, the second largest element swapped with the second smallest element, etc.  The flipping operator will be denoted by $\mathcal{F}$.  
\end{definition}

@@

\emph{Yajit wrote this section.}

@@

\emph{Deepak wrote this section.}

@@

\begin{theorem}
\label{rsk}
There exists an algorithm (called the \emph{RSK algorithm}) which defines a bijection between the set of permutations in $S(n)$ avoiding a decreasing subsequence of length 3 and the set of pairs of size-$n$ SYT of the same shape and at most two rows. 
\end{theorem}\todo{give a citation for this}
Using all this machinery and all these results, we are finally able to prove the closed-form expression for $s_n(321)$.
\begin{theorem}
$s_n(321)$ and $s_n(123)$ are equal to $C_n$, the $n^{th}$ Catalan number.
\end{theorem}
\begin{proof}
By the reversing lemma we only need to prove this result for $s_n(321)$. By Theorem \ref{rsk}, there are $s_n(321)$ elements in the set $A$ of pairs of size-$n$ SYT of the same shape and at most two rows. By Theorem \ref{catalan_2}, the set $B_n$ of SYT of shape $(n, n)$ has its size given by $C_n$. We will construct a bijection between the sets $A_n$ and $B_n$: because $|A_n|=s_n(321)$ and $|B_n|=C_n$, it will therefore follow that $s_n(321)=C_n$, as desired.

@@

\emph{Deepak wrote this section. Leon and Yajit closely edited this section.}

>>>>

\emph{Yajit wrote this section. Leon closely edited this section.}

@@

\subsection{Flipping}
In this section we introduce the concept of flipping to prove the `Flipping Lemma.'
\begin{definition}
We define the \emph{flip} of a sequence $a$ as the sequence $b$ with the same elements as $a$, but with the largest element swapped with the smallest element, the second largest element swapped with the second smallest element, etc.  The flipping operator will be denoted by $\mathcal{F}$.  
\end{definition}

@@

\emph{Yajit wrote this section. Leon edited this section.}

@@

\emph{Deepak wrote this section. Leon closely edited this section.}

@@

\begin{theorem}\footnote{A more detailed exposition of this result can be found at http://math.mit.edu/~jnovak/d.18.312/lecture16.pdf. The result itself appears on page 9.}
\label{rsk}
There exists an algorithm (called the \emph{RSK algorithm}) which defines a bijection between the set of permutations in $S(n)$ avoiding a decreasing subsequence of length 3 and the set of pairs of size-$n$ SYT of the same shape and at most two rows. 
\end{theorem}
Using all this machinery and all these results, we are finally able to prove the closed-form expression for $s_n(321)$.
\begin{theorem}
$s_n(321)$ and $s_n(123)$, the number of permutations on $n$ elements which avoid 321 and 123 respectively, are equal to $C_n$, the $n^{th}$ Catalan number.
\end{theorem}
\begin{proof}
By the reversing lemma we only need to prove this result for $s_n(321)$. By Theorem \ref{rsk}, there are $s_n(321)$ elements in the set $A$ of pairs of size-$n$ SYT of the same shape and at most two rows. By Theorem \ref{catalan_2}, the set $B_n$ of SYT of shape $(n, n)$ has its size given by $C_n$. We will construct a bijection between the sets $A_n$ and $B_n$: because $|A_n|=s_n(321)$ and $|B_n|=C_n$, it will therefore follow that $s_n(321)=C_n$, as desired.

@@

\emph{Deepak wrote this section. Leon rewrote parts of it. Yajit closely edited this section.}
